Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-10-18 Manuscript/manuscript.tex

\begin{table}
  \caption{The relative surface excess and the probabilities of contacts between bound Calcium and different
    oxygen moieties in POPC from different simulation models. The CaCl$_2$ concentration of the ions in water
    is 350mM in the analyzed systems\label{tab:binding} }
  \begin{tabular}{l|c c c}
    model                 & $\Gamma_{Ca}^w$ (nm$^{-2}$)   & $P^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $P^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
    \hline
    ECC-lipids            &  $0.07 \pm 0.01 $            &  99\%   &    25\%    \\
    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
    Lipid14/Dang          &  $ 0.3 \pm 0.03  $           &  -        &    -       \\
  \end{tabular}
\end{table}
To quantify the binding affinities in different models we calculated 
the relative surface excess from Eq.~\ref{surfexcess}. 
Calcium cations at a molar concentration of $350\,\mathrm{mM}$ provide
$\Gamma_i^w = 0.07 \pm 0.01 \rm{nm}^{-2}$ in the ECC-lipid model, which is significantly
smaller than the values for the Lipid14 model with \AA{}qvist ions, $\Gamma_i^w = 0.13 \pm 0.01 \rm{nm}^{-2}$,
or with Dang ions, $\Gamma_i^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
\todo{JOE: I'd put the $\Gamma$ values into a table SAMULI: They are now in the table \label{tab:binding} 
with contacts of calcium with different oxygens. However, I am not sure if there is enough information for the
table unless we analyze the contacts for Lipid14 models as well.}
Interestingly, the relative surface excess of 
\ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) is qualitatively different from \ce{CaCl2}
having a negative value $\Gamma_i^w = -0.1 \pm 0.01 \rm{nm}^{-2}$, 
meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface. 
This is in contrast to the most of the available lipid force fields,
which which predict a specific binding of sodium on 
PC lipid bilayers \cite{catte16}.
%\todo{JOE: Add references to NaCl simulation papers that show peaks.
%Add a simple note that there is NBFIX trick in Charmm36 (?).} 
%SAMULI: I think that it is enough to cite NMRlipids II here, which shows all the results
This finding is depicted in Figures~\ref{fig:nacl-dens} and~\ref{fig:cacl-dens},
which show the density profile of ions and water along the membrane normal. 
%The density profiles show
%a larger Ca$^{2+}$ density peak in lipid headgroup region for
%the Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
\todo{Densities and relative surface excess from NaCl systems to be added.
  The discussion below to be finished after this.}

>>>>

\begin{table}
  \caption{The relative surface excess and the probabilities of contacts between bound Calcium and different
    oxygen moieties in POPC from different simulation models. The CaCl$_2$ concentration of the ions in water
    is 350mM in the analyzed systems\label{tab:binding} }
  \begin{tabular}{l|c c c}
    model                 & $\Gamma_{Ca}^w$ (nm$^{-2}$)   & $P^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $P^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
    \hline
    ECC-lipids            &  $0.07 \pm 0.01 $            &  99\%   &    25\%    \\
    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
    Lipid14/Dang          &  $ 0.3 \pm 0.03  $           &  -        &    -       \\
  \end{tabular}
\end{table}
The binding affinities of Ca$^{2+}$ ions to POPC bilayer in different simulation 
models were quantified by calculating the relative surface excess using Eq.~\ref{surfexcess} 
and the bulk concentrations determined from the density profiles in Fig.~\ref{fig:cacl-dens}. 
All these simulation have the same, $350\,\mathrm{mM}$, concentration of CaCl$_2$ with 
respect to water (concentration before solvating the lipids in experiments).
As expected from the changes of the lipid headgroup order parameters in Fig. \ref{fig:delta_ordPar_CaCl}, 
the relative surface excess for the ECC-lipid model, $\Gamma_{\rm Ca}^w = 0.07 \pm 0.01 \rm{nm}^{-2}$,
is significantly smaller than for the Lipid14 model with \AA{}qvist ions, $\Gamma_{\rm Ca}^w = 0.13 \pm 0.01 \rm{nm}^{-2}$,
or with Dang ions, $\Gamma_{\rm Ca}^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
\todo{JOE: I'd put the $\Gamma$ values into a table SAMULI: They are now in the table \label{tab:binding} 
together with probabilities of the calcium binding probabilities to different POPC oxygens. 
However, I am not sure if there the table is useful in the end.}
Interestingly, the relative surface excess of 
\ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) 
calculated using density profiles in Fig.~\ref{fig:nacl-dens}
is qualitatively different from \ce{CaCl2}
having a negative value $\Gamma_i^w = -0.1 \pm 0.01 \rm{nm}^{-2}$, 
meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface. 
This is in contrast to the most of the available lipid force fields,
which predict a specific binding of sodium on 
PC lipid bilayers \cite{catte16}.
%\todo{JOE: Add references to NaCl simulation papers that show peaks.
%Add a simple note that there is NBFIX trick in Charmm36 (?).} 
%SAMULI: I think that it is enough to cite NMRlipids II here, which shows all the results
%
%This finding is depicted in Figures~\ref{fig:nacl-dens} and~\ref{fig:cacl-dens},
%which show the density profile of ions and water along the membrane normal. 
%
%The density profiles show
%a larger Ca$^{2+}$ density peak in lipid headgroup region for
%the Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
\todo{Densities and relative surface excess from NaCl systems to be added.
  The discussion above to be finished after this.}
